---
title: "PCA"
output: html_document
date: '2022-04-28'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Data loading
```{r}
my_data_clean_aug <- read_tsv(file = "data/03_my_data_clean_aug.tsv",show_col_types = FALSE)
```

#Data preprocessing for PCA
```{r}
PCA_data <- my_data_clean_aug %>% 
  select(where(is.numeric),
        -(matches("energy|Log|eff|time.sec|power"))) %>% 
  as_tibble()
```



##one hot encoding with "caret" package

#joining data (population, sex)
```{r}
one_hot_enc_data <- my_data_clean_aug %>% 
  select(c("Population",
           "Sex"))
```

```{r}
one_hot_enc_data_join <- PCA_data %>% 
  mutate(one_hot_enc_data$Population, 
         one_hot_enc_data$Sex)
```

#renaming columns
```{r}
one_hot_enc_data_join <- one_hot_enc_data_join %>% 
  rename( "Population" = "one_hot_enc_data$Population", 
          "Sex" = "one_hot_enc_data$Sex")

```

#Creatin dummies (caret package)
```{r}
#define one-hot encoding function
dummy <- dummyVars(" ~ .",
                   data=one_hot_enc_data_join)

#perform one-hot encoding on data frame
final_data <- data.frame(predict(dummy,
                                 newdata=one_hot_enc_data_join))

#view final data frame
final_data

```

#PCA with one hot encoding
```{r}
pca_fit <- final_data %>% 
  prcomp(scale = TRUE)
```

#Eigenvalues
```{r}
pca_fit %>%
  tidy(matrix = "eigenvalues")
```

#Eigenvalues percentage plot
```{r}
pca_fit %>%
  tidy(matrix = "eigenvalues") %>%
  ggplot(aes(PC, percent)) +
  geom_col(fill = "#56B4E9",
           alpha = 0.8) +
  scale_x_continuous(breaks = 1:9) +
  scale_y_continuous(labels = scales::percent_format(),
    expand = expansion(mult = c(0,
                                0.01))) 
```
#Contribution of variable to PC
```{r}
var<-get_pca_var(pca_fit)
a<-fviz_contrib(pca_fit,
                "var",
                axes=1,
                xtickslab.rt=90) # default angle=45°
plot(a, main = "Variables percentage contribution of first Principal Components")
```


#PC1 VS PC2 Plot - Population
```{r}
pl1 <- pca_fit %>%
  augment(my_data_clean_aug) %>% # add original dataset back in
  ggplot(aes(.fittedPC1,
             .fittedPC2,
             color = Population)) + 
  geom_point(size = 1.5)

pl2 <- pca_fit %>%
  augment(my_data_clean_aug) %>% # add original dataset back in
  ggplot(aes(.fittedPC1,
             .fittedPC2,
             color = Sex)) + 
  geom_point(size = 1.5)
  
pl3 <- pca_fit %>%
  augment(my_data_clean_aug) %>% # add original dataset back in
  ggplot(aes(.fittedPC1,
             .fittedPC2,
             color = distance_class)) + 
  geom_point(size = 1.5)

show(pl1 / pl3 + pl2)
```

#Rotation matrix
```{r}
pca_fit %>%
  tidy(matrix = "rotation")
```

#Plot rotation matrix(arrows)
```{r}
# define arrow style for plotting
arrow_style <- arrow(
  angle = 20, ends = "first",
  type = "closed",
  length = grid::unit(8, "pt")
)

# plot rotation matrix
pca_fit %>%
  tidy(matrix = "rotation") %>%
  pivot_wider(names_from = "PC",
              names_prefix = "PC",
              values_from = "value") %>%
  ggplot(aes(PC1,
             PC2)) +
  geom_segment(xend = 0,
               yend = 0,
               arrow = arrow_style) +
  geom_text(aes(label = column),
    hjust = 1,
    nudge_x = -0.02, 
    color = "#904C2F") +
  xlim(-1.25, .5) +
  ylim(-.5, 1) +
  coord_fixed()  # fix aspect ratio to 1:1
 



```

##K N N


##4. Clustering Principal Components


```{r}
kmean <- pca_fit$x %>%
  kmeans(centers = 2,
         iter.max = 1000,
         nstart = 10) %>%
  augment(final_data)


```

```{r}
pl4 <- pca_fit %>%
  augment(kmean) %>% # add original dataset back in
  ggplot(aes(.fittedPC1,
             .fittedPC2,
             color = .cluster)) + 
  geom_point(size = 1.5)

show(pl4)
```


